---
title: "Practical Machine Learning Final Project"
author: "crusanton"
date: "11/9/2017"
output: html_document
---

## Project Introduction
Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. 

## Data Locations
Training:  https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

Test:      https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

### Load Libraries
```{r LibraryChunk, echo=TRUE,cache= TRUE}
#opts_chunk$set(echo = TRUE, results = 'hold')
library(caret)
library(gbm)
library(knitr)
library(randomForest)
library(RColorBrewer)
library(rpart)
library(rpart.plot)
```

### Upload Datasets and Partition
```{r UploadChunk, echo=TRUE,cache= TRUE}
set.seed(12345)
TrainingURL <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
TrainingDataset <- read.csv(url(TrainingURL), na.strings=c("NA","#DIV/0!",""))
TestingDataset <- read.csv(url(testUrl), na.strings=c("NA","#DIV/0!",""))
inTrain <- createDataPartition(TrainingDataset$classe, p=0.6, list=FALSE)
myTraining <- TrainingDataset[inTrain, ]
myTesting <- TrainingDataset[-inTrain, ]
# Check size of Training 
dim(myTraining)
# Check size of Testing
dim(myTesting)
```

### Clean Data
```{r CleanChunk, echo=TRUE,cache= TRUE}
# Remove the Near Zero Variance
HoldNZV <- nearZeroVar(myTraining, saveMetrics=TRUE)
myTraining <- myTraining[,HoldNZV$nzv==FALSE]
HoldNZV<- nearZeroVar(myTesting,saveMetrics=TRUE)
myTesting <- myTesting[,HoldNZV$nzv==FALSE]
myTraining <- myTraining[c(-1)]

HoldTraining <- myTraining
for(i in 1:length(myTraining)) {
    if( sum( is.na( myTraining[, i] ) ) /nrow(myTraining) >= .7) {
        for(j in 1:length(HoldTraining)) {
            if( length( grep(names(myTraining[i]), names(HoldTraining)[j]) ) == 1)  {
                HoldTraining <- HoldTraining[ , -j]
            }   
        } 
    }
}

# Get rid of unnecessary columns
myTraining <- HoldTraining
rm(HoldTraining)
ClassWith <- colnames(myTraining)
ClassWithout <- colnames(myTraining[, -58])      # remove classe 
myTesting <- myTesting[ClassWith]                # compare columns
TestingDataset <- TestingDataset[ClassWithout]   # compare columns

# Same data type
for (i in 1:length(TestingDataset) ) {
    for(j in 1:length(myTraining)) {
        if( length( grep(names(myTraining[i]), names(TestingDataset)[j]) ) == 1)  {
            class(TestingDataset[j]) <- class(myTraining[i])
        }      
    }      
}

TestingDataset <- rbind(myTraining[2, -58] , TestingDataset)
TestingDataset <- TestingDataset[-1,]
```

### Confusion Matrix
```{r ConfusionChunk, echo=TRUE,cache= TRUE}
set.seed(12345)
Model1 <- rpart(classe ~ ., data=myTraining, method="class")

CMPrediction <- predict(Model1, myTesting, type = "class")
CMTreeData <- confusionMatrix(CMPrediction, myTesting$classe)
CMTreeData
                                       
plot(CMTreeData$table, col = CMTreeData$byClass, main = paste("Decision Tree Confusion Matrix: Accuracy =", round(CMTreeData$overall['Accuracy'], 4)))
```

### Random Tree
```{r RandomChunk, echo=TRUE,cache= TRUE}
set.seed(12345)
Model2 <- randomForest(classe ~ ., data=myTraining)
RTPrediction <- predict(Model2, myTesting, type = "class")
CMRFdata <- confusionMatrix(RTPrediction, myTesting$classe)
CMRFdata

plot(Model2)
plot(CMRFdata$table, col = CMTreeData$byClass, main = paste("Random Forest Confusion Matrix: Accuracy =", round(CMRFdata$overall['Accuracy'], 4)))
```

### Generalized Boosted Regression (GBM)
```{r GBMChunk, echo=TRUE,cache= TRUE}
set.seed(12345)
TrainControlModel <- trainControl(method = "repeatedcv", number = 5, repeats = 1)
TrainGBM <- train(classe ~ ., data=myTraining, method = "gbm", trControl = TrainControlModel, verbose = FALSE)
GBMPrediction <- predict(TrainGBM, newdata=myTesting)
GBMTesting <- confusionMatrix(GBMPrediction, myTesting$classe)
GBMTesting
plot(TrainGBM, ylim=c(0.9, 1))
```

### Predict on Testing
```{r PredictChunk, echo=TRUE,cache= TRUE}
TestingPrediction <- predict(Model2, TestingDataset, type = "class")
TestingPrediction
```

### Write to file
```{r WriteChunk, echo=TRUE,cache= TRUE}
WritePrediction = function(x){
    n = length(x)
    for(i in 1:n){
        filename = paste0("problem_id_",i,".txt")
        write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
    }
}

WritePrediction(TestingPrediction)

```